---
title: "top_100_heatmap_and_indicative_species_analyze"
author: "Jingzhe Jiang"
date: "2015年10月19日"
output: 
  html_document:
    toc: true
    toc_depth: 4
    keep_md: true
---
# top 100 otus
combine top 100 otus of each oyster samples in cohort 2

## read data
```{r}
library(dplyr)
origin_data <- read.csv("D:/bioinfo/UBC_data/otu_summaries/otu_table_os_top960_NormAndRanked.csv", header = TRUE, row.names = 1) # single number giving the column of the table which contains the row names. But dpyr never preserve row names, you should use add_rownames() to preserve them!
str(origin_data)
row.names(origin_data) %>% head
colnames(origin_data)
mydata <- origin_data %>% add_rownames %>% tbl_df
mydata %>% glimpse
row.names(mydata) %>% head
```
## select top 100 otus ids
from each samples (os7-os12), than combined and get rid of redundences, than filtrate from my data
```{r}
top_7 <- mydata %>% 
    arrange(desc(os7)) %>% 
      select(rowname) %>% 
        head(100)
top_8 <- mydata %>% 
    arrange(desc(os8)) %>% 
      select(rowname) %>% 
        head(100)
top_9 <- mydata %>% 
    arrange(desc(os9)) %>% 
      select(rowname) %>% 
        head(100)
top_10 <- mydata %>% 
    arrange(desc(os10)) %>% 
      select(rowname) %>% 
        head(100)
top_11 <- mydata %>% 
    arrange(desc(os11)) %>% 
      select(rowname) %>% 
        head(100)
top_12 <- mydata %>% 
    arrange(desc(os12)) %>% 
      select(rowname) %>% 
        head(100)
# all = TRUE means full Outer join tables
top_cohort2 <- merge(top_12,top_11, by = "rowname", all = TRUE)
top_cohort2 <- merge(top_cohort2,top_10, by = "rowname", all = TRUE)
top_cohort2 <- merge(top_cohort2,top_9, by = "rowname", all = TRUE)
top_cohort2 <- merge(top_cohort2,top_8, by = "rowname", all = TRUE)
top_cohort2 <- merge(top_cohort2,top_7, by = "rowname", all = TRUE)
top_cohort2_data <- merge(mydata, top_cohort2, by = "rowname", all.y = TRUE)
top_cohort2_data %>% arrange(desc(os8)) %>% head
```
## normalized columns
based on the total reads of each sample
```{r}
# clip total reads from excel, than paste into R
refs <- read.delim("refs.txt", header = TRUE)
refs %>% glimpse
## top_normal <- top_cohort2_data[,2:13]*311167/as.numeric(refs[1,2:13])
## 两个dataframe相除必须维度一致，或者将分母转变为vector才可以相除。但above is not correct, but I don't know why!
top_normal <- top_cohort2_data[,2:13]*311167/refs[rep(1,251),2:13]
#两个dataframe相除必须维度一致，因此用rep(1,251)复制了refs中的ttlReads数据，使其变成dataframe。参见如下：
refs[rep(1,5),1:13]
rownames(top_normal) <- top_cohort2_data$rowname
top_normal %>% str
top_cohort2_data %>% str
identical(row.names(top_normal),top_cohort2_data$rowname)
identical(top_normal$os2,top_cohort2_data$os2) # one is num, one is int
identical(as.integer(top_normal$os2),top_cohort2_data$os2)
```
## heatmap with ggplot2
```{r}
library(ggplot2)
library(reshape2)
# the dataframe is converted from wide format to a long format
top_melt <- top_normal %>% 
  add_rownames %>% 
    select(rowname,os7,os8,os9,os10,os11,os12) %>% 
      melt
(p <- ggplot(top_melt, aes(variable, rowname)) + geom_tile(aes(fill = log10(value)), colour = "white") + scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 2) + theme(axis.text.y = element_text(size = 1), axis.text.x = element_text(size = 4)) + coord_fixed(ratio=0.15))
```
## cluster otus on heatmap
```{r}
library(vegan)
den_order <- top_normal[,10:11] %>% 
  vegdist(method = "bray") %>% 
    hclust(method = "average") %>% 
      as.dendrogram 

table <- top_normal %>% add_rownames

table$rowname <- table$rowname[order.dendrogram(den_order)]
# > nba$Name <- with(nba, reorder(Name, PTS))


top_melt <- table %>% 
    select(rowname,os7,os8,os9,os10,os11,os12) %>% 
      melt
(p <- ggplot(top_melt, aes(variable, rowname)) + geom_tile(aes(fill = log10(value)), colour = "white") + scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 2) + theme(axis.text.y = element_text(size = 1), axis.text.x = element_text(size = 4)) + coord_fixed(ratio=0.15))
```
## replot with different methods
above result is clustered not very good. So I tried with different subset data (line 92) or dist methods (line 93). But results are all similar. 
then i use reorder() plot based on the value of os10.
```{r}
os10 <- table %>% 
    select(rowname,os10)
table$rowname <- with(os10,reorder(rowname, os10))
# > nba$Name <- with(nba, reorder(Name, PTS))

top_melt <- table %>% 
    select(rowname,os7,os8,os9,os10,os11,os12) %>% 
      melt
(p <- ggplot(top_melt, aes(variable, rowname)) + geom_tile(aes(fill = log10(value)), colour = "white") + scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 2) + theme(axis.text.y = element_text(size = 1), axis.text.x = element_text(size = 4)) + coord_fixed(ratio=0.15))
```
## pickout high abundant otus (>=500) from os10 and os11
just like the first analyz do.
```{r}
top_10 <- top_cohort2_data %>% 
    arrange(desc(os10)) %>% 
      select(rowname, os10) %>% 
        filter(os10>=500)
plot(log10(top_10$os10))
top_11 <- top_cohort2_data %>% 
    arrange(desc(os11)) %>% 
      select(rowname, os11) %>% 
        filter(os11>=500)
plot(log10(top_11$os11))
top_cohort2 <- merge(top_10,top_11, by = "rowname", all = TRUE) # finally there 86 otus were kept.
top_cohort2_data <- merge(mydata, top_cohort2, by = "rowname", all.y = TRUE)
top_normal <- top_cohort2_data[,2:13]*311167/refs[rep(1,86),2:13]
rownames(top_normal) <- top_cohort2_data$rowname
identical(row.names(top_normal),top_cohort2_data$rowname)
identical(as.integer(top_normal$os2),top_cohort2_data$os2)
den_order <- top_normal[,7:12] %>% 
  vegdist(method = "bray") %>% 
    hclust(method = "average") %>% 
      as.dendrogram 
table <- top_normal %>% add_rownames
table$rowname <- table$rowname[order.dendrogram(den_order)]
# > nba$Name <- with(nba, reorder(Name, PTS))


top_melt <- table %>% 
    select(rowname,os7,os8,os9,os10.x,os11.x,os12) %>% 
      melt
(p <- ggplot(top_melt, aes(variable, rowname)) + geom_tile(aes(fill = log10(value)), colour = "white") + scale_fill_gradient2(low = "green", mid = "yellow", high = "red", midpoint = 2.2) + theme(axis.text.y = element_text(size = 1), axis.text.x = element_text(size = 4)) + coord_fixed(ratio=0.15))

```
It looks like total different from first analysis. So decide to load first data to compare with each other
```{r}
firstdata <- read.csv("otu_table_os_200otu--s0806-0813otu_Normed.csv", header=TRUE,row.names=1, sep=",")
firstdata["denovo976096",1:6]/top_normal["denovo976096",7:12]
firstdata["denovo208223",1:6]/top_normal["denovo208223",7:12]
firstdata["denovo976096",1:6]/top_normal["denovo976096",7:12]
# have tried different rows. the ratio is always around 2.5. Except a few otus was absent in one of these two datas.
```
plot firstdata with ggplot
```{r}
den_order <- firstdata[,1:6] %>% 
  vegdist(method = "bray") %>% 
    hclust(method = "average") %>% 
      as.dendrogram 
table <- firstdata %>% add_rownames

table$rowname <- table$rowname[order.dendrogram(den_order)]
top_melt <- table %>% 
    select(rowname,h0730,h0806,h0813,s0806,s0813,s0820) %>% 
      melt
(p <- ggplot(top_melt, aes(variable, rowname)) + geom_tile(aes(fill = log10(value)), colour = "white") + scale_fill_gradient2(low = "green", mid = "yellow", high = "red", midpoint = 3) + theme(axis.text.y = element_text(size = 1), axis.text.x = element_text(size = 4)) + coord_fixed(ratio=0.15))
```
I found it is also different from first result. I found the row names were changed but the value doesn't changed.
that because this commond: **`table$rowname <- table$rowname[order.dendrogram(den_order)]`**
please check for below, you will know why!
`table1 <- firstdata %>% add_rownames`
`table2 <- firstdata %>% add_rownames`
`table1$rowname <- table$rowname[63:1]`
`table2$rowname <- table$rowname[1:63]`
then I just true `with()` method to reorder rows, but still different from first result
```{r}
table <- firstdata %>% add_rownames
table$den_order <- order.dendrogram(den_order)
table$rowname <- with(table, reorder(rowname,den_order))
top_melt <- table %>% 
    select(rowname,h0730,h0806,h0813,s0806,s0813,s0820) %>% 
      melt
(p <- ggplot(top_melt, aes(variable, rowname)) + geom_tile(aes(fill = log10(value)), colour = "white") + scale_fill_gradient2(low = "green", mid = "yellow", high = "red", midpoint = 3) + theme(axis.text.y = element_text(size = 1), axis.text.x = element_text(size = 4)) + coord_fixed(ratio=0.15))
```
Though otus and their value has been reordered, but the output of heatmap is still different from the first analysis. This maybe because `with()` and `reorder()` method is not suit for dendrogram heatmap plot. I still should learn from template heatmap to plot a dendrogram heatmap!
## So Plot with `Heatplus`
```{r}
library(Heatplus)
library(RColorBrewer)
scalegreenred <- colorRampPalette(c("green", "red"), space = "rgb")(100)
heatmap(as.matrix(firstdata[,1:6]), Rowv = den_order, Colv = NA, col = scalegreenred, cexRow = 0.6) 
```
YES! it works! then plot with other subset data
### pickout high abundant otus (>=500) from os10 and os11
```{r}
den_order <- top_normal[,7:12] %>% 
  vegdist(method = "bray") %>% 
    hclust(method = "average") %>% 
      as.dendrogram 
heatmap(as.matrix(top_normal), Rowv = den_order, Colv = NA, col = scalegreenred, cexRow = 0.6) 
```
## find indicative species based on mortality data
```{r}
library(indicspecies)
data(wetland)
# transform mydata based on wetland data
tr_data <- top_normal %>% t %>% data.frame
# check transformed data
tr_data[,"denovo1017076"]
top_normal["denovo1017076",]
# 设定2种可能的分类标准，根据mortality data
classF1 <- c("a","a","a","d","a","a","a","a","a","b","d","a")
classF2 <- c("a","a","a","d","b","a","a","a","a","b","d","a")
# run from classF1 to classF6
indval = multipatt(tr_data, classF1, control = how(nperm=999))
summary(indval)
indval = multipatt(tr_data, classF2, control = how(nperm=999))
summary(indval)
```
only denovo705886 has significancy with group b+d. It's: denovo705886	3.145617504	0	4.413126107	11.91115449	0	0	10.5581394	14.72814299	0	**577.5020538**	12.52276834	3.366478779	o-Rickettsiales; f-mitochondria
## find indicative species based on 960 otus data
```{r}
firstall <- read.csv("otu_table_os_200otu_NormAndRanked.csv", header=TRUE,row.names=1, sep=",")
tr_data <- firstall[,1:12] %>% t %>% data.frame
indval = multipatt(tr_data, classF1, control = how(nperm=999))
summary(indval)
indval = multipatt(tr_data, classF2, control = how(nperm=999))
summary(indval)
```
For classF1: Group b+d  #sps.  14 
               stat p.value    
denovo274616  0.998   0.001 ***
denovo705886  0.997   0.004 ** 
denovo1038783 0.993   0.008 ** 
denovo1503755 0.971   0.006 ** 

For classF2: Group b+d  #sps.  10 
               stat p.value   
denovo1699894 0.995   0.004 **
denovo656814  0.994   0.002 **
denovo230815  0.992   0.008 **
denovo1205230 0.991   0.006 **
denovo314706  0.980   0.003 **

